---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipewelder

**pipewelder** is an R package for streamlined retrieval and cleaning of economic and public health time series data from:

- **Bureau of Labor Statistics (BLS)**
- **Federal Reserve Economic Data (FRED)**
- **Centers for Disease Control and Prevention (CDC)**

It consolidates API interactions into consistent, high-level pipeline functions:

- `get_bls()` for labor force data
- `get_fred()` for market and macroeconomic indicators
- `get_cdc()` for mortality and health surveillance data

## Installation

You can install the development version of pipewelder from [GitHub](https://github.com/velvetvaldes/pipewelder) with:

```r
# install.packages("pak")
pak::pak("velvetvaldes/pipewelder")
```

## Examples

### Retrieve labor force data from BLS

```{r bls-example, eval = FALSE}
# Set your API key
Sys.setenv(BLS_KEY = "your_api_key")

# Retrieve unemployment data from 2000 to 2023
bls_df <- pipewelder::get_bls("LNS14000000", 2000, 2023)
head(bls_df)
```

### Retrieve market data from FRED

```{r fred-example, eval = FALSE}
fred_df <- pipewelder::get_fred("VIXCLS", "1990-01-01")
head(fred_df)
```

### Retrieve health surveillance data from CDC

```{r cdc-example, eval = FALSE}
cdc_df <- pipewelder::get_cdc()
head(cdc_df)
```

---

Generated from `README.Rmd` using `devtools::build_readme()`.
