---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pipewelder

**pipewelder** is an R package for streamlined retrieval and cleaning of economic and public health time series data from:

- **Bureau of Labor Statistics (BLS)**
- **Federal Reserve Economic Data (FRED)**
- **Centers for Disease Control and Prevention (CDC)**

It provides two layers of functionality:

- The `get_*()` functions retrieve data for a single series ID from each respective source.
- The `construct_*()` functions operate on a named list of series metadata tibbles (such as those returned by `get_full_gsheet()`), iterating over each entry to retrieve and clean data for multiple series at once.

The output is a named list of tibbles, where each tibble represents a cleaned time series. Instead of inserting the series ID as a column, each tibble is tagged using `attr(x, "series_id")` to preserve the series identifier without modifying the data structure.

Functions included:

- `get_bls()` retrieves BLS labor force data
- `get_fred()` retrieves FRED macroeconomic indicators
- `get_cdc()` retrieves CDC mortality and health surveillance data
- `construct_bls()` builds cleaned BLS data from a list of series metadata
- `construct_fred()` builds cleaned FRED data from a list of series metadata
- `construct_cdc()` builds cleaned CDC data from a list of metadata tabs, auto-detecting schema structure

Every `construct_*()` function has a caching mechanism which caches the retrieved data to `~\pipewelder-cache`.  The cache can be flushed with the `flush = TRUE` parameter but should retrieve fresh data once each calendar day.  All `construct_*()` functions also include default date parameters, which differ by source. Be sure to review the documentation to understand the behavior for each.

## Installation

You can install `pipewelder` from [GitHub](https://github.com/velvet-valdes/pipewelder) with:

```{r installation-example, eval= FALSE}
install.packages("pak") # Install 'pak' if not present
pak::pak("velvet-valdes/pipewelder")
```

## Examples

### Retrieve a Google sheet to use as a input for the `construct_*()` functions:

```{r  gsheet-example, eval = FALSE}
url <- "https://docs.google.com/spreadsheets/d/your-sheet-id/edit#gid=0"
sheet <- get_full_gsheet(url)
```

### Retrieve series data from BLS:

```{r bls-example, eval = FALSE}
# Set your API key
Sys.setenv(BLS_KEY = "your_api_key")

# Retrieve a single series of BLS data from 2000 to 2023
bls_df <- get_bls("LNS14000000", 2000, 2023)

# Construct a list of tibbles using a Google Sheet as input
BLS <- construct_bls(sheet)
```

### Retrieve series data from FRED
```{r fred-example, eval = FALSE}
# Set your API key
Sys.setenv(FRED_KEY = "your_api_key")

# Retrieve a single series of FRED data from 1990 to the present
fred_df <- get_fred("VIXCLS", "1990-01-01")

# Construct a list of tibbles using a Google Sheet as input
FRED <- construct_fred(sheet)
```

### Retrieve series data from CDC

```{r cdc-example, eval = FALSE}
# Retrieve a single series of CDC data with date defaults
cdc_df <- get_cdc("9j2v-jamp")

# Construct a list of tibbles using a Google Sheet as input
CDC <- construct_cdc(sheet)
```

---

Generated from `README.Rmd` using `devtools::build_readme()`.
